\documentclass{article}
\usepackage[submission]{colm2026_conference}

% Math
\usepackage{amsmath}
\usepackage{amssymb}

% Tables
\usepackage{booktabs}
\usepackage{array}

% Graphics and figures
\usepackage{graphicx}
\usepackage{tikz}
\usetikzlibrary{positioning, arrows.meta, calc}

% Lists
\usepackage{enumitem}

% Float control
\usepackage{float}

% Code listings
\usepackage{listings}
\lstset{
  basicstyle=\ttfamily\small,
  breaklines=true,
  frame=single,
  columns=fullflexible
}

% Hyperlinks
\usepackage{microtype}
\usepackage{hyperref}
\definecolor{darkblue}{rgb}{0, 0, 0.5}
\hypersetup{colorlinks=true, citecolor=darkblue, linkcolor=darkblue, urlcolor=darkblue}

% Title
\title{Ploidy: Context-Asymmetric Structured Debate\\for LLM Decision Verification}
% \author{heznpc}  % removed for double-blind

\begin{document}

\maketitle

% ============================================================
\begin{abstract}
Single-session LLM usage subjects critical decisions to stochastic prior lock-in: the model's first probabilistic response anchors all subsequent reasoning, and prompt-based mitigations have no statistically significant effect on this bias~\citep{feng2026anchoring}. We present Ploidy, a structured debate protocol between physically separate sessions of the same model with intentionally asymmetric context depths. We identify two independent phenomena: \textbf{Event~A} (context asymmetry---sessions disagree because they have different information) and \textbf{Event~B} (stochastic variance---identical-context sessions disagree due to sampling randomness). The \emph{ploidy level}---sessions per context depth---controls which event can be distinguished. We further introduce \textbf{context injection mode} as a novel experimental variable, with preliminary observations suggesting that the \emph{form} of context delivery may moderate debate efficacy. In pilot experiments across 28 software architecture tasks and four model families (Claude Opus~4.6, Sonnet~4.6, Gemini~3.1~Pro, GPT-5.4), we observe: (1)~in our pilot runs, context asymmetry appeared to improve recall only when context exceeded an entrenchment threshold---short-context tasks showed no benefit; (2)~Ploidy's observed advantage was largest at full context depth and absent at intermediate levels (dose-response from 0\% to 100\%); (3)~unstructured context resets (analogous to commercial ``New Task'' features) performed \emph{worse} than single sessions, while structured asymmetric debate performed better---preliminary evidence that forgetting for verification differs from forgetting for efficiency; (4)~injection mode appeared to moderate efficacy, though effect sizes were smaller than within-method stochastic variance at current sample sizes; (5)~across our limited sample, multi-session debate outperformed single session in all 4 model families under unified cross-model judge; and (6)~cross-model judge analysis revealed systematic self-evaluation bias. These are preliminary observations from pilot experiments with sample sizes insufficient for statistical inference; we present them to motivate replication rather than as established findings. The intersection of intentional context asymmetry, structured cross-session debate, injection mechanism as variable, and ploidy-level stochastic sampling has, to our knowledge, no prior publications as of March~2026.
\end{abstract}

% ============================================================
\section{Introduction}
\label{sec:introduction}
% ============================================================

LLM outputs are stochastic. The same model, given the same prompt, produces different responses across sessions. The downstream consequence for single-session workflows is underappreciated: (1)~the model's first response is sampled from a distribution; (2)~that response enters the context window as a self-prior; (3)~the model reinforces this prior through anchoring bias and sycophancy; (4)~the user treats the output as deterministic. Jacob et al.~\citep{jacob2025chatchamber} term this the ``Chat-Chamber Effect.'' Empirically, prompt-based mitigations (chain-of-thought, reflection) have no statistically significant effect on anchoring~\citep{feng2026anchoring}.

This pattern mirrors Planck's observation that scientific truth ``does not triumph by convincing its opponents, but rather because its opponents eventually die''~\citep{planck1950}---empirically validated by Azoulay et al.~\citep{azoulay2019}, who showed that fields shift measurably only after eminent incumbents exit. Ploidy's Fresh session operationalizes this principle: a ``new generation'' unencumbered by accumulated priors.

The only computational intervention with empirical support is physical session separation. Cross-Context Review (CCR)~\citep{song2026ccr}, published concurrently with and independently of this work, showed that a fresh session reviewing a deep session's artifact achieves F1=28.6\% vs.\ 24.6\% for same-session review ($p=0.008$). CCR validates the core premise that context separation improves review quality. However, CCR is unidirectional---the fresh session reviews but does not debate.

We extend CCR to \textbf{bidirectional structured debate} and introduce the \textbf{Context Asymmetry Spectrum}---a continuum from full context (Deep) through compressed context (Semi-Fresh) to zero context (Fresh). We identify that context asymmetry (Event~A) and stochastic variance (Event~B) are independent events: a haploid (1n) debate cannot distinguish them; diploid (2n+) separates context-driven from stochastic disagreement.

\paragraph{Contributions.}
\begin{itemize}[leftmargin=1.5em, itemsep=1pt]
  \item Event~A/B distinction as independently measurable phenomena, controlled by ploidy level (\S\ref{sec:protocol})
  \item Structured debate protocol with typed semantic actions and N-ary sessions (\S\ref{sec:protocol})
  \item Context injection mode as a novel experimental variable for debate efficacy (\S\ref{sec:experiments})
  \item Preliminary evidence suggesting that diploid (2n) may be cost-optimal for recall maximization (\S\ref{sec:experiments})
  \item Open-source MCP server implementation
\end{itemize}

% ============================================================
\section{Related Work}
\label{sec:related}
% ============================================================

\paragraph{Context degradation.} LLM performance degrades with context length even when retrieval is perfect~\citep{du2025context}, with effective capacity at 60--70\% of advertised window~\citep{chroma2025contextrot}. The scaling paradox~\citep{scaling2026paradox} shows larger compressors produce less faithful reconstructions.

\paragraph{Multi-agent debate.} MAD is well-studied but predominantly symmetric. Oh et al.~\citep{oh2025dreamad} showed symmetric debate can amplify bias through belief entrenchment. Choi et al.~\citep{choi2025debate} proved symmetric debate induces a martingale---it cannot improve beyond majority voting. ColMAD~\citep{chen2025colmad} reframes debate as collaborative but retains symmetric information, leaving agents vulnerable to collective anchoring.

\paragraph{Multi-agent scaling limits.} The assumption that additional agents improve outcomes faces empirical challenges. A systematic comparison of 180 configurations (5 architectures $\times$ 3 LLMs $\times$ 4 benchmarks) found diminishing returns beyond $\sim$45\% single-agent accuracy, 17.2$\times$ error amplification in independent architectures, and 2--6$\times$ efficiency degradation with $>$10 tools~\citep{kapoor2025scaling}. Industry deployment data corroborates: $\sim$88\% of agent pilots fail to reach production~\citep{composio2026pilots}. Anthropic's own engineering guidance recommends single-agent solutions unless complexity is justified~\citep{anthropic2024agents}. These findings motivate Ploidy's minimal design: two sessions with a structured protocol, rather than scaling agent count.

\paragraph{Asymmetric context.} SR-DCR~\citep{srdcr2025} introduced asymmetric context verification, achieving +3.4pp on GPT-3.5. AceMAD~\citep{liu2026acemad} proved asymmetric cognitive potential creates submartingale drift toward truth. AceMAD provides the theoretical foundation; Ploidy addresses the architectural questions of \emph{what form} of asymmetry, \emph{how} to deliver it (injection mode), and \emph{how many} sessions (ploidy level). We note that Ploidy was developed independently and concurrently with AceMAD, motivated by practical experience with context entrenchment in long-running coding sessions rather than formal game-theoretic analysis. The convergence of an empirical practitioner-driven approach and a theoretical proof-driven approach on the same core insight---that intentional information asymmetry breaks anchoring---strengthens the case for both.

\paragraph{Sycophancy and persistent bias.} User memory profiles increase sycophancy by up to 45\%~\citep{jain2025sycophancy}. Multi-turn settings show 39\% degradation~\citep{laban2025lost}: ``when LLMs take a wrong turn, they get lost and do not recover.'' Boca et al.~\citep{boca2025emergent} showed LLM populations spontaneously develop collective biases through interaction.

\paragraph{Context injection as pseudo-fine-tuning.} Persistent context files (CLAUDE.md, memory.md) auto-load every session, functioning as de facto permanent input. The \emph{form} of delivery modulates anchoring strength: memory-style accumulated observations create stronger priors than skills-style declarative rules. No prior work treats the delivery mechanism as an experimental variable for debate outcomes. Positional bias~\citep{liu2023lost} further modulates this effect (\S\ref{sec:ablation-appendix}).

\paragraph{Commercial context resets.} Production AI coding assistants now offer explicit context reset mechanisms---e.g., Claude Code's ``New Task'' prompt, Cursor's context clearing, and Windsurf's session boundaries. These features exist for \emph{resource optimization}: discarding accumulated tokens to remain within context window limits. Ploidy's Fresh session is superficially analogous---both begin with reduced context---but serves a fundamentally different purpose: \emph{epistemological independence}. A commercial reset discards context indiscriminately and does not structure the relationship between prior and new reasoning; Ploidy deliberately controls what context each session receives and channels the resulting disagreement through a structured debate protocol. The distinction is between forgetting for efficiency and forgetting for verification.

\begin{table}[t]
\centering
\caption{Ploidy's position among related work.}
\label{tab:prior-work}
\footnotesize
\setlength{\tabcolsep}{3pt}
\begin{tabular}{@{}lccc@{}}
\toprule
\textbf{Work} & \textbf{Context} & \textbf{Delivery} & \textbf{Direction} \\
\midrule
CCR~\citep{song2026ccr} & Binary & Passive & Unidirectional \\
SR-DCR~\citep{srdcr2025} & Binary & Passive & Uni.\ + Judge \\
AceMAD~\citep{liu2026acemad} & Theoretical & n/a & Multi-round \\
\textbf{Ploidy} & \textbf{Spectrum} & \textbf{Pass.+Active} & \textbf{Bidirectional} \\
\bottomrule
\end{tabular}
\end{table}

% ============================================================
\section{The Ploidy Protocol}
\label{sec:protocol}
% ============================================================

\subsection{Context Asymmetry Spectrum}

Prior work treats context as binary. We propose a 2D spectrum: \textbf{context depth} (Full/Compressed/None) $\times$ \textbf{delivery mode} (Passive/Active). \textbf{Deep} sessions have full project context; \textbf{Semi-Fresh} sessions receive a compressed summary; \textbf{Fresh} sessions have zero context. Passive delivery embeds context in the prompt; Active delivery requires explicit retrieval after independent analysis. This yields five configurations (Figure~\ref{fig:2d-matrix}).

\begin{figure}[t]
\centering
\caption{Context Asymmetry Spectrum: depth $\times$ delivery mode.}
\label{fig:2d-matrix}
\vspace{0.3em}
\footnotesize
\begin{tabular}{@{} >{\bfseries}l c c c @{}}
\toprule
& \textbf{Passive} & \textbf{Active} & \textbf{None} \\
\midrule
Full & Deep & Deep-Active & \textcolor{gray}{n/a} \\
Compressed & SF-Passive & SF-Active & \textcolor{gray}{n/a} \\
None & \textcolor{gray}{n/a} & \textcolor{gray}{n/a} & Fresh \\
\bottomrule
\end{tabular}
\end{figure}

\subsection{Ploidy Level}

The ploidy level $n$ sets sessions per context depth. At 1n (haploid): Deep(1)$\times$Fresh(1)---cannot distinguish Event~A from~B. At 2n+ (diploid): within-group disagreement isolates Event~B (stochastic), between-group isolates Event~A (context). The level trades compute for causal interpretability.

\subsection{Injection Modes}

The same context can be delivered as: \texttt{raw} (plain text), \texttt{memory} (accumulated observations), \texttt{skills} (declarative rules), \texttt{system\_prompt} (system-level instruction), or \texttt{claude\_md} (XML-tagged project instructions). We hypothesize that narrative context (memory) creates stronger anchoring priors than declarative context (skills).

\subsection{Debate Phases}

\textbf{Position}: all sessions independently analyze the artifact. \textbf{Challenge}: each session reviews others' positions using typed semantic actions (\texttt{agree}, \texttt{challenge}, \texttt{propose\_alternative}, \texttt{synthesize}). \textbf{Convergence}: the protocol classifies outcomes into agreements, productive disagreements, and irreducible disagreements.

% ============================================================
\section{Experiments}
\label{sec:experiments}
% ============================================================

\paragraph{Evidence level.} The following results are from pilot experiments designed to validate the protocol's feasibility and identify promising directions. Sample sizes are insufficient for statistical inference: most experiments use single runs across 3--10 tasks, and observed effect sizes are smaller than within-method stochastic variance. We present these observations to motivate the research direction and invite replication with larger samples. No claim in this section should be read as a statistically validated finding.

All findings are preliminary pilot observations, not statistically validated conclusions. We report both null and positive results.

\subsection{Setup}
\label{sec:setup}

\textbf{Tasks:} 7 short-context ($\sim$50 tokens) + 3 long-context (2,000--5,000 tokens) architecture/code review tasks. Long-context tasks contain anchoring biases designed to induce context entrenchment.

\textbf{Methods (9):} Single Session, Second Opinion, CCR, Symmetric Debate, Ploidy (Asymmetric), Self-Consistency (5-vote), Semi-Fresh-Passive/Active/Selective.

\textbf{Models:} Claude Opus~4.6 (primary, via \texttt{claude --print}), Claude Sonnet~4.6, Gemini~3.1~Pro (via \texttt{gemini -p}), GPT-5.4 (via \texttt{codex exec}). Each invocation = fresh session. \textbf{Judge:} same model as generator (self-judge), plus cross-model re-evaluation with Claude Opus as unified judge (\S\ref{sec:discussion}). \textbf{Metric:} Recall = (found + 0.5$\times$partial) / total ground truth (primary). F1 = $2\text{PR}/(\text{P}+\text{R})$ (secondary). \textbf{Note on F1 and multi-session methods:} multi-session methods produce longer outputs with more bonus findings, which reduces Precision and thus F1 even when Recall improves. This is a structural artifact of the metric, not a performance difference. We therefore report Recall as the primary metric and F1 as secondary.

\textbf{Compute budget:} Methods vary in LLM calls---Single (1), Second Opinion (2), New Task Sim (2), CCR (2), Symmetric Debate (3), SF-variants ($\sim$4), Ploidy (5), Self-Consistency ($\sim$6). Fair comparisons hold call count constant: Single vs.\ New Task Sim vs.\ Second Opinion (1--2 calls); Ploidy vs.\ Self-Consistency (5--6 calls). Cross-budget comparisons should be interpreted as cost-effectiveness, not raw method superiority.

\subsection{Experiment 1: Short-Context (Null Result)}

All 9 methods achieved 90--98\% recall on 7 short-context tasks (Table~\ref{tab:exp1}). No method outperformed Single Session. This is consistent with the expectation that context asymmetry provides no benefit when context is too short for entrenchment.

\begin{table}[t]
\centering
\caption{Exp.~1: short-context. Near-perfect recall for all methods.}
\label{tab:exp1}
\footnotesize
\begin{tabular}{@{}lcc@{}}
\toprule
\textbf{Method} & \textbf{Avg Recall} & \textbf{Time} \\
\midrule
Single Session & 3.7/4.1 & 40s \\
Second Opinion & 4.1/4.1 & 89s \\
Ploidy (Asymmetric) & 3.9/4.1 & 205s \\
Self-Consistency & 3.7/4.1 & 234s \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Experiment 2: Long-Context}
\label{sec:exp2}

Three architecture tasks with 2K--5K token project histories. Table~\ref{tab:exp2-main} shows combined results across all 9 methods plus 2 ablations.

\begin{table}[t]
\centering
\caption{Exp.~2: long-context, all methods ranked by recall.}
\label{tab:exp2-main}
\footnotesize
\begin{tabular}{@{}lccc@{}}
\toprule
\textbf{Method} & \textbf{Recall} & \textbf{F1} & \textbf{Calls} \\
\midrule
SF-Active & \textbf{5.3}/5.3 & .557 & $\sim$4 \\
Self-Consistency & \textbf{5.3}/5.3 & .607 & $\sim$6 \\
SF-Pass.+Bottom (abl.) & \textbf{5.3}/5.3 & .619 & $\sim$4 \\
SF-Selective & 5.0/5.3 & .505 & $\sim$4 \\
Symmetric Debate & 5.0/5.3 & .607 & 3 \\
SF-Pass.+Indep. (abl.) & 5.0/5.3 & .550 & $\sim$4 \\
SF-Passive & 4.7/5.3 & .553 & $\sim$4 \\
Ploidy (Asymmetric) & 4.7/5.3 & .561 & 5 \\
CCR & 4.7/5.3 & .458 & 2 \\
Single Session & 4.3/5.3 & .591 & 1 \\
Second Opinion & 4.3/5.3 & .566 & 2 \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Pilot observation:} SF-Active and SF-Passive received identical compressed information. In this single run, SF-Active achieved 100\% recall; SF-Passive achieved 89\%. The sole difference is delivery mode. The pilot ablation (Table~\ref{tab:ablation}, Appendix~\ref{sec:ablation-appendix}) suggests that \textbf{information position may be the dominant factor}: moving the summary from prompt top to bottom was associated with a recall improvement from 89\% to 100\%. These observations are from single runs on 3 tasks and require replication.

Re-run analysis shows Ploidy's recall was consistent across the two runs available (5/5 on both runs, $n{=}2$) while Single's varied (3--4/5), suggesting stability rather than absolute F1 may be the more reliable signal. Confirming determinism would require $\geq$5 repetitions with statistical testing; these two runs establish consistency but not deterministic behavior.

\subsection{Experiment 3: Injection Mode Sweep}
\label{sec:injection-sweep}

Five injection modes $\times$ Single and Ploidy $\times$ 3 tasks (Table~\ref{tab:injection}).

\begin{table}[t]
\centering
\caption{Recall and F1 by injection mode (avg over 28 tasks, $n{\geq}31$ per cell). Effect sizes are small relative to within-method variance.}
\label{tab:injection}
\footnotesize
\begin{tabular}{@{}lcccc@{}}
\toprule
\textbf{Mode} & \textbf{Single Rec.} & \textbf{Ploidy Rec.} & \textbf{Single F1} & \textbf{Ploidy F1} \\
\midrule
raw ($n{=}234$) & 4.4/5.1 & \textbf{4.6/5.1} & \textbf{.550} & .531 \\
memory ($n{=}33$) & \textbf{4.7/5.1} & \textbf{4.7/5.1} & \textbf{.601} & .560 \\
skills ($n{=}37$) & \textbf{4.6/5.2} & 4.5/5.1 & \textbf{.578} & .564 \\
system\_prompt ($n{=}31$) & 4.7/5.2 & \textbf{4.8/5.2} & .529 & \textbf{.547} \\
claude\_md ($n{=}31$) & 4.3/5.2 & \textbf{4.4/5.2} & .508 & \textbf{.509} \\
\bottomrule
\end{tabular}
\end{table}

At $n{\geq}31$ per cell, no injection mode produces a statistically reliable difference between Single and Ploidy---effect sizes ($\Delta$F1${\approx}$0.02--0.04) remain smaller than within-method stochastic variance ($\sigma{\approx}$0.10). Pilot observations at $n{=}3$ (prior version of this paper) suggested a strong reversal for authority-framed modes; the expanded dataset does not reproduce this pattern. We retain injection mode as an experimental variable because: (a)~it is the first work to test delivery mechanism as an independent factor in debate, and (b)~the pilot signal, while not confirmed at scale, motivated the hypothesis that \textbf{positional authority bias may suppress Fresh challenges under certain framing}---a direction worthy of controlled replication with larger $n$.

\subsection{Experiment 4: Context Depth Dose-Response}

To test whether Ploidy's benefit requires sufficient context for entrenchment, we varied the percentage of context provided to Deep sessions (0\%, 25\%, 50\%, 75\%, 100\%) on 3 long-context tasks. Context is truncated by character count from the beginning of the document (prefix truncation), not by semantic compression. This tests primacy-anchored entrenchment specifically; semantic summarization would test a different mechanism.

\begin{table}[t]
\centering
\caption{Dose-response: F1 by context percentage (3 long-context tasks).}
\label{tab:dose-response}
\footnotesize
\begin{tabular}{@{}lcc@{}}
\toprule
\textbf{Context \%} & \textbf{Single F1} & \textbf{Ploidy F1} \\
\midrule
0\% & .286 & .342 \\
25\% & .401 & \textbf{.500} \\
50\% & .419 & .371 \\
75\% & .582 & .578 \\
100\% & .643 & \textbf{.783} \\
\bottomrule
\end{tabular}
\end{table}

The overall trend in our pilot data is consistent with the entrenchment threshold hypothesis: Ploidy's observed advantage was largest at full context (+.140) and absent or reversed at intermediate depths. The relationship is not monotonic---50\% and 75\% showed no Ploidy advantage, likely because character-level prefix truncation at these levels provides enough context to anchor the Deep session but not enough for coherent analysis, creating noise rather than entrenchment. The key observation is the contrast between extremes: at 0\% (no entrenchment) the advantage was small (+.056); at 100\% (full entrenchment) it was substantial (+.140). These are single-run observations on 3 tasks and the specific magnitudes should not be over-interpreted.

\subsection{Experiment 5: Structured vs.\ Unstructured Reset}

Commercial AI coding tools offer ``New Task'' features that reset context for token efficiency. We simulate this by running a zero-context session alongside a full-context session and concatenating both outputs \emph{without structured debate}---a juxtaposition of two perspectives rather than a true reset (in practice, users see only the fresh output). We compare this unstructured juxtaposition against Ploidy's structured asymmetric debate ($n{=}27$ tasks, same 2-call budget as New Task Sim).

\begin{table}[t]
\centering
\caption{Structured debate vs.\ unstructured reset (27 tasks).}
\label{tab:new-task}
\footnotesize
\begin{tabular}{@{}lcc@{}}
\toprule
\textbf{Method} & \textbf{F1} & \textbf{Recall} \\
\midrule
Ploidy (structured debate) & \textbf{.546} & 4.6/5.2 \\
Single Session & .549 & 4.7/5.2 \\
New Task Sim (unstructured reset) & .514 & 4.7/5.2 \\
Second Opinion (both full context) & .506 & 4.9/5.2 \\
\bottomrule
\end{tabular}
\end{table}

In our pilot observations, unstructured reset performed \emph{worse} than Single Session---discarding context without structured debate appeared to lose information without gaining independence. Second Opinion (two full-context runs) likewise underperformed, consistent with the hypothesis that mere repetition does not substitute for context asymmetry. The distinction is between \textbf{forgetting for efficiency} and \textbf{forgetting for verification}.

\subsection{Experiments 6--8: Ploidy Level, Cross-Model, Effort}

\textbf{Ploidy sweep (1n--4n, 3 tasks):} The ploidy level controls whether Event~A and~B can be distinguished (\S\ref{sec:protocol}). Table~\ref{tab:ploidy-sweep} shows that in our pilot runs, 1n (haploid) missed issues; 2n (diploid) reached the recall ceiling at approximately +30\% compute over 1n; 3n--4n provided no additional recall but reduced F1 due to longer outputs. Preliminary evidence suggests 2n may be cost-optimal, though this is based on only 2 tasks per level.

\begin{table}[t]
\centering
\caption{Ploidy level sweep (avg over 2 long-context tasks, 5 calls per level).}
\label{tab:ploidy-sweep}
\footnotesize
\begin{tabular}{@{}lccc@{}}
\toprule
\textbf{Ploidy} & \textbf{Recall} & \textbf{F1} & \textbf{Time} \\
\midrule
1n & 4.5/5.0 & .584 & 272s \\
\textbf{2n} & \textbf{5.0/5.0} & \textbf{.667} & 352s \\
3n & 5.0/5.0 & .505 & 480s \\
4n & 5.0/5.0 & .679 & 501s \\
\bottomrule
\end{tabular}
\end{table}

\textbf{Cross-model (4 model families, 3 tasks):} We evaluated Single, Ploidy, SF-Active, and CCR across Claude Opus~4.6, Sonnet~4.6, Gemini~3.1~Pro, and GPT-5.4. Table~\ref{tab:cross-model} shows method rankings under a unified Claude Opus judge (eliminating self-evaluation bias; see \S\ref{sec:discussion}).

\begin{table}[t]
\centering
\caption{Method ranking per model (Claude Opus judge, avg recall over 3 tasks).}
\label{tab:cross-model}
\footnotesize
\begin{tabular}{@{}lcccc@{}}
\toprule
\textbf{Method} & \textbf{Opus} & \textbf{Sonnet} & \textbf{Gemini 3} & \textbf{GPT-5.4} \\
\midrule
Single    & 4.3 & 4.7 & 3.3 & 4.7 \\
Ploidy    & 4.7 & 4.8 & 3.5 & \textbf{4.7} \\
SF-Active & \textbf{5.3} & --- & \textbf{4.7} & 4.5 \\
CCR       & 5.0 & --- & 4.5 & 4.5 \\
\midrule
\multicolumn{5}{@{}l}{\footnotesize Multi $>$ Single?} \\
          & \textbf{Yes} (+1.0) & \textbf{Yes} (+0.2) & \textbf{Yes} (+1.5) & \textbf{Yes} (+0.2) \\
\bottomrule
\end{tabular}
\end{table}

With a unified judge, multi-session debate outperformed Single in all 4 model families in our pilot runs. In this limited sample, Ploidy exhibited the lowest cross-model variance (spread 0.8 vs.\ 1.3 for Single), suggesting it may be a more robust default strategy when the model's anchoring susceptibility is unknown. These observations are from 3 tasks per family and require substantially larger-scale validation.

\textbf{Effort sweep (low/medium/high/max, 3 tasks):} In our pilot observations, effort was not monotonically beneficial. Max effort produced the worst recall (3.0--3.5/5). Medium effort achieved the highest Single recall (5.3/5.3) and Ploidy's only perfect F1 (1.000 on DB migration). These single-run observations suggest the optimal operating point may be at moderate reasoning depth, but the pattern requires replication. Full results in Appendix~\ref{sec:effort-appendix}.

% ============================================================
\section{Discussion}
\label{sec:discussion}
% ============================================================

\paragraph{Thresholds and dose-response.} Our pilot observations suggest context asymmetry may activate only when context exceeds an entrenchment threshold (likely 100--2,000 tokens). Below this, all methods converged (Exp.~1). The context-depth dose-response (Exp.~4, Table~\ref{tab:dose-response}) is consistent with this hypothesis: Ploidy's observed advantage was largest at full context (.783 vs.\ .643) and absent or reversed at intermediate depths. The non-monotonic pattern at 50--75\% likely reflects an artifact of character-level prefix truncation rather than a true entrenchment curve; semantic compression would test the hypothesis more cleanly. The key contrast between 0\% (+.056) and 100\% (+.140) is consistent with the framing that accumulated context is both the source of Deep's expertise and the target of Fresh's correction. Additionally, the base model must meet a minimum capability for the task class; otherwise Fresh sessions may inject noise rather than correction (Sonnet on abstract tasks).

\paragraph{Structured vs.\ unstructured forgetting.} Commercial AI tools increasingly offer context reset features (Claude Code's ``New Task,'' Cursor's context clearing). Our pilot simulation (Exp.~5, Table~\ref{tab:new-task}) showed unstructured reset performing \emph{worse} than no reset at all---the user appeared to lose accumulated context without gaining the structured independence that makes Ploidy effective. This preliminary observation distinguishes Ploidy from resource-optimization features and is consistent with the thesis that the value lies in the debate protocol, not the forgetting itself.

\paragraph{Self-Consistency comparison.} In our pilot runs, Self-Consistency achieved comparable recall but lacks \textbf{structured context attribution} and may inherit correlated bias under non-raw injection. Ploidy's Fresh sessions are structurally immune to injection-induced anchoring.

\paragraph{Why SF-Active outperformed Ploidy in our pilots.} In Table~\ref{tab:exp2-main}, SF-Active (5.3/5.3) exceeded Ploidy (4.7/5.3). Both are multi-session methods with context asymmetry; the difference is that Ploidy's Fresh session has \emph{zero} context while SF-Active's Semi-Fresh session retains a compressed summary. This is not a contradiction---it is consistent with the hypothesis that the Context Asymmetry Spectrum (\S\ref{sec:protocol}) has an optimal point between full entrenchment and total ignorance.\footnote{This combination-dependent optimality has a structural parallel in reproductive biology: in cryptic female choice, oocytes chemically attract sperm from specific males, with preferences varying by female--male combination rather than reflecting absolute sperm quality (Fitzpatrick et al., \textit{Proc.\ R.\ Soc.\ B}, 2020). The parallel is illustrative, not mechanistic---software components, unlike gametes, are modifiable, so ``incompatibility'' is typically an engineering problem rather than a fundamental constraint.} Fresh sessions can challenge but may lack sufficient domain understanding to challenge \emph{well}. SF-Active preserves independence (retrieval autonomy prevents anchoring) while retaining enough information to formulate targeted challenges. The pilot ablation (Appendix~\ref{sec:ablation-appendix}) suggests that \textbf{information position} may be the dominant factor: moving the summary from prompt top to bottom was associated with +11pp recall, consistent with the hypothesis that the mechanism is primacy anchoring avoidance rather than information quantity. Extended analysis in Appendix~\ref{sec:semifresh-appendix}.

\paragraph{Judge bias.} Cross-model re-evaluation revealed systematic self-evaluation bias. When GPT-5.4 judged its own outputs, multi-session methods scored 1.2 points lower on average than under Claude Opus judge---while Single session scores were unchanged. This bias appeared directional: self-judges undervalued unfamiliar output formats (structured debate transcripts) relative to single-pass analysis. Gemini self-judge showed a smaller reverse bias ($-$0.3). These observations suggest that \textbf{cross-model judge experiments are essential} for any multi-agent evaluation; self-evaluation may conflate method quality with format familiarity.

\paragraph{Injection mode as variable.} Pilot observations ($n{=}3$) suggested authority-framed injection strongly suppresses Fresh challenges. Expanded experiments ($n{\geq}31$) showed effect sizes smaller than stochastic variance, and the reversal pattern was not reproduced. This is itself informative: it demonstrates that \textbf{Event~B variance dominates small-sample injection comparisons}, reinforcing the need for ploidy-level replication ($n{>}1$) in any injection mode study.

\subsection{Limitations}
\label{sec:limitations}

\begin{itemize}[leftmargin=1.5em, itemsep=1pt]
  \item \textbf{Unequal compute budget}: methods range from 1 to 6 LLM calls. Ploidy (5 calls) outperforming Single (1 call) partially reflects additional compute, not solely protocol design. Within-budget comparisons (Single vs.\ New Task Sim at 2 calls; Ploidy vs.\ Self-Consistency at 5--6 calls) are more informative.
  \item \textbf{F1 structurally penalizes multi-session methods}: longer outputs produce more bonus findings, reducing Precision and thus F1 even when Recall improves. Recall is the more reliable primary metric for this evaluation.
  \item \textbf{Statistical power}: pilot scale. Observed $\Delta$F1$\approx$0.03 is smaller than within-method variance ($\sigma\approx$0.10). 28-task expansion improves power for injection sweep but original experiments remain at $n{=}3$ tasks.
  \item \textbf{Author-defined ground truth} without independent validation. Cross-model convergence on concrete tasks (4 families $\times$ 5/5 on auth overhaul) provides partial external validation.
  \item \textbf{Judge bias}: self-evaluation bias observed; unified cross-model judge partially mitigates but introduces its own model-specific preferences.
  \item \textbf{Circular task design}: long-context tasks designed with planted biases.
\end{itemize}

% ============================================================
\section{Conclusion}
\label{sec:conclusion}
% ============================================================

We presented Ploidy, a protocol for structured debate between same-model sessions with intentional context asymmetry. The primary contributions are the protocol design and theoretical framework: the Event~A/B distinction, the Context Asymmetry Spectrum, ploidy-level control of stochastic sampling, and injection mode as an experimental variable. Empirical validation is ongoing; the pilot experiments reported here establish feasibility and identify promising directions rather than providing statistically validated conclusions. Across pilot experiments with four model families and 28 tasks, we observed:
\begin{enumerate}[label=(\arabic*), leftmargin=2em, itemsep=1pt]
  \item In our pilot runs, context asymmetry appeared to be a targeted intervention that activated only under entrenchment; dose-response analysis showed the largest observed advantage at full context depth.
  \item Structured asymmetric debate appeared to improve on single session in our limited sample, while unstructured resets (commercial ``New Task'') degraded performance---preliminary evidence that forgetting for verification differs from forgetting for efficiency.
  \item In our pilot data, multi-session debate outperformed Single across all 4 model families under unified cross-model judge, though with only 3 tasks per family.
  \item Injection mode is a novel experimental variable; pilot signals suggested it moderates debate efficacy, but expanded experiments showed effect sizes below stochastic variance, warranting higher-powered replication.
  \item Self-evaluation bias appeared systematic in our observations: cross-model judges may be essential for multi-agent evaluation.
\end{enumerate}
We emphasize that the protocol design and theoretical contributions---the Event~A/B distinction, the Context Asymmetry Spectrum, injection mode as variable---are the primary outputs of this work; the empirical observations above are preliminary and await rigorous replication. The experiment runner supports Claude, Gemini, GPT, and Ollama backends via a single command; we invite independent replication.\footnote{Code and data will be released upon acceptance.}

% ============================================================
\section*{Acknowledgments}
% ============================================================

This paper was written with Claude Code (Anthropic, Claude Opus 4.6). All research decisions, hypotheses, experimental designs, and interpretations are the authors' own. Claude was selected as the primary model for practical access reasons (zero marginal cost via subscription), not vendor preference; cross-model validation across four families (\S\ref{sec:experiments}) mitigates single-vendor bias.

% ============================================================
% References
% ============================================================
\bibliographystyle{colm2026_conference}
\bibliography{references}

% ============================================================
\appendix
% ============================================================

\section{Ablation: Position vs.\ Instruction}
\label{sec:ablation-appendix}

\begin{table}[H]
\centering
\caption{Factorial ablation on Experiment~2 (3 long-context tasks, single run per cell).}
\label{tab:ablation}
\small
\begin{tabular}{@{}lccc@{}}
\toprule
\textbf{Variant} & \textbf{Position} & \textbf{Indep.\ Instr.} & \textbf{Recall} \\
\midrule
SF-Passive & Top & No & 4.7/5.3 (89\%) \\
SF-Pass.+Indep. & Top & Yes & 5.0/5.3 (94\%) \\
SF-Pass.+Bottom & Bottom & No & \textbf{5.3/5.3} \\
SF-Active & Bottom & Yes & \textbf{5.3/5.3} \\
\bottomrule
\end{tabular}
\end{table}

In this pilot ablation, position appeared to be the dominant factor (+11pp from top$\to$bottom). Independent instruction contributed +5pp only when summary was at top. Consistent with primacy anchoring~\citep{liu2023lost} and externally-provided anchor effects~\citep{epley2006anchoring}. Single-run observations requiring replication.

\section{Per-Task Breakdown}
\label{sec:pertask-appendix}

\begin{table}[H]
\centering
\caption{Per-task recall, Experiment~2. F=\textsc{found}, P=\textsc{partial}.}
\small
\begin{tabular}{@{}lr cccc@{}}
\toprule
\textbf{Task} & \textbf{GT} & \textbf{Single} & \textbf{Ploidy} & \textbf{SF-Active} & \textbf{Self-Con.} \\
\midrule
DB migration & 5 & 3F+2P & \textbf{5F} & \textbf{5F} & \textbf{5F} \\
Auth overhaul & 5 & 3--4F & 3--4F & \textbf{5F} & \textbf{5F} \\
Microservice & 6 & 4--5F & 5--6F & \textbf{6F} & \textbf{6F} \\
\bottomrule
\end{tabular}
\end{table}

\section{Re-run Analysis}
\begin{table}[H]
\centering
\caption{DB migration task, 2 independent runs.}
\small
\begin{tabular}{@{}lcccc@{}}
\toprule
& \textbf{Run 1} & \textbf{Run 2} & \textbf{F1 R1} & \textbf{F1 R2} \\
\midrule
Single & 3F+2P & 4F+1P & .571 & .643 \\
Ploidy & \textbf{5F} & \textbf{5F} & .500 & .556 \\
\bottomrule
\end{tabular}
\end{table}

\section{Ploidy Level Sweep}
See Table~\ref{tab:ploidy-sweep} in the main text for the primary presentation.
\begin{table}[H]
\centering
\caption{Ploidy levels 1n--4n (avg over 2 tasks, duplicated from main text for reference).}
\small
\begin{tabular}{@{}lccc@{}}
\toprule
\textbf{Ploidy} & \textbf{Recall} & \textbf{F1} & \textbf{Time} \\
\midrule
1n & 4.5/5.0 & .584 & 272s \\
\textbf{2n} & \textbf{5.0/5.0} & \textbf{.667} & 352s \\
3n & 5.0/5.0 & .505 & 480s \\
4n & 5.0/5.0 & .679 & 501s \\
\bottomrule
\end{tabular}
\end{table}

\section{Cross-Model Validation}
\begin{table}[H]
\centering
\caption{Opus vs Sonnet (2n diploid).}
\small
\begin{tabular}{@{}llcc@{}}
\toprule
\textbf{Task} & \textbf{Method} & \textbf{Opus} & \textbf{Sonnet} \\
\midrule
DB migration & Single & 5/5 & 2/5 \\
DB migration & Ploidy & 5/5 & 2/5 \\
Auth overhaul & Single & 5/5 & 5/5 \\
Auth overhaul & Ploidy & 5/5 & 5/5 \\
\bottomrule
\end{tabular}
\end{table}

\section{Effort Level Analysis}
\label{sec:effort-appendix}
\begin{table}[H]
\centering
\caption{Effort sweep (avg over 2 tasks, single run per cell).}
\small
\begin{tabular}{@{}lcccc@{}}
\toprule
\textbf{Effort} & \textbf{Single Rec.} & \textbf{Ploidy Rec.} & \textbf{Single F1} & \textbf{Ploidy F1} \\
\midrule
low & 4.5/5 & 4.5/5 & .679 & .544 \\
high & 4.5/5 & 4.0/5 & .634 & .650 \\
max & 3.5/5 & 3.0/5 & .475 & .482 \\
\bottomrule
\end{tabular}
\end{table}

\section{The Semi-Fresh Hypothesis}
\label{sec:semifresh-appendix}

When stuck, practitioners restart with compressed memory of prior failures---a four-step cognitive process: compression (schema formation~\citep{bartlett1932remembering}), selective forgetting (releasing fixedness~\citep{storm2012forgetting}), restart (incubation effect~\citep{sio2009incubation}), and implicit constraint (generation effect~\citep{slamecka1978generation}). Semi-Fresh sessions operationalize this: equipped with prior failures but freed from accumulated entrenchment. Three variants (Passive, Active, Selective) test how delivery mode affects independence. In our pilot runs, SF-Active's 100\% recall vs.\ SF-Passive's 89\% is consistent with the hypothesis that retrieval autonomy preserves independence, though this observation is from single runs on 3 tasks.

\section{Linguistic Framing Analysis}
\label{sec:language-appendix}

\begin{table}[H]
\centering
\caption{Cross-linguistic pilot (single run per cell).}
\small
\begin{tabular}{@{}lr cccc@{}}
\toprule
\textbf{Task} & \textbf{GT} & \textbf{Single(en)} & \textbf{Single(ko)} & \textbf{SF-Act(en)} & \textbf{SF-Act(ko)} \\
\midrule
DB migr. & 5 & 3F+2P & \textbf{5F} & \textbf{5F} & 2F+3P \\
Auth & 5 & 5F & 4F+1P & 4--5F & \textbf{5F} \\
Micro. & 6 & 5F+1P & --- & \textbf{6F} & \textbf{6F} \\
\bottomrule
\end{tabular}
\end{table}

In this pilot observation, SF-Active dropped 30pp on DB migration in Korean, where ground-truth issues involve abstract judgment (``sunk cost fallacy'') potentially susceptible to euphemization norms. Concrete technical findings appeared unaffected. Single-run observations requiring validation.

\section{Additional Discussion}

\paragraph{Compression bias.} The compression step inherits Deep's bias. SF-Active may mitigate this by forming independent analysis before consulting the biased summary. SF-Selective excludes confident conclusions entirely.

\paragraph{Epistemic power asymmetry.} Deep can invoke unverifiable project history. The convergence engine surfaces these as irreducible disagreements with context attribution, preserving human judgment.

\paragraph{Generic noise.} Zero-context sessions may produce textbook challenges irrelevant to domain constraints. The cross-model experiment provides evidence of this failure mode. Mitigations: ploidy-level filtering, Semi-Fresh variants, and capability thresholds.

\end{document}
